% bib/bibliography-uk.tex — «Список бібліографічного опису» у форматі ДСТУ 8302:2015.
% Порядок відповідає порядку цитування у тексті (стиль unsrt).

\begin{citreflist}

\item Anthropic.
  Model Context Protocol.
  \textit{modelcontextprotocol.io}.
  URL: \url{https://modelcontextprotocol.io}
  (дата звернення: 01.05.2026).

\item Anisimov A.~V., Derevianchenko O.~V., Kuliabko P.~P., Fedorus O.~M.
  Programming System PARCS.
  \textit{Journal of Computer and Communications}.
  2017. Vol.~5, No.~9. С.~129--139.
  DOI: \url{https://doi.org/10.4236/jcc.2017.59009}.

\item Bohusevych O.
  PARCS --- A .NET-Based Distributed Computing Platform on Kubernetes.
  GitHub. 2026.
  URL: \url{https://github.com/alexeybogusevich/parcs8}
  (дата звернення: 01.05.2026).

\item Bohusevych O., Derevianchenko O.
  Building Highly Scalable Parallel Compute Systems with PARCS, Kubernetes and KEDA.
  \textit{Bulletin of Taras Shevchenko National University of Kyiv. Physical and Mathematical Sciences}.
  2026.

\item Chaturvedi A., Nichols D., Singh S., Bhatele A.
  HPC-Coder-V2: Studying Code LLMs Across Low-Resource Parallel Languages.
  arXiv:2412.15178. 2024.
  URL: \url{https://arxiv.org/abs/2412.15178}.

\item Hasan M.~M., Li H., Rajbahadur G.~K., Adams B., Hassan A.~E.
  Model Context Protocol (MCP) Tool Descriptions Are Smelly!
  Towards Improving AI Agent Efficiency with Augmented MCP Tool Descriptions.
  arXiv:2602.14878. 2026.
  URL: \url{https://arxiv.org/abs/2602.14878}.

\item Hou X., Zhao Y., Wang S., Wang H.
  Model Context Protocol (MCP): Landscape, Security Threats, and Future Research Directions.
  arXiv:2503.23278. 2025.
  URL: \url{https://arxiv.org/abs/2503.23278}.

\item Kim S., Moon S., Tabrizi R., Lee N., Mahoney M.~W., Keutzer K., Gholami A.
  An LLM Compiler for Parallel Function Calling.
  \textit{Proc. 41st International Conference on Machine Learning (ICML 2024)}.
  PMLR, 2024.
  С.~24370--24391.

\item Ma Y., Gou Z., Hao J., Xu R., Wang S., Pan L., Yang Y., Cao Y., Sun A.
  SciAgent: Tool-Augmented Language Models for Scientific Reasoning.
  \textit{Proc. 2024 Conference on Empirical Methods in Natural Language Processing (EMNLP 2024)}.
  Association for Computational Linguistics, 2024.
  С.~15701--15736.

\item Mialon G., Dessì R., Lomeli M., Nalmpantis C., Pasunuru R.,
  Raileanu R. та ін.
  Augmented Language Models: A Survey.
  \textit{Transactions on Machine Learning Research}. 2023.

\item Nader N., Diehl P., Brandt S., Kaiser H.
  LLM \& HPC: Benchmarking DeepSeek's Performance in High-Performance Computing Tasks.
  \textit{Proc. ISC High Performance 2025}.
  Springer, 2025.
  (Lecture Notes in Computer Science).
  arXiv:2504.03665.

\item Nichols D., Davis J.~H., Xie Z., Rajaram A., Bhatele A.
  Can Large Language Models Write Parallel Code?
  \textit{Proc. 33rd International Symposium on High-Performance Parallel and Distributed Computing (HPDC 2024)}.
  ACM, 2024.
  С.~117--130.
  URL: \url{https://doi.org/10.1145/3625549.3658689}.

\item Patil S.~G., Zhang T., Wang X., Gonzalez J.~E.
  Gorilla: Large Language Model Connected with Massive APIs.
  arXiv:2305.15334. 2023.
  URL: \url{https://arxiv.org/abs/2305.15334}.

\item Qin Y., Liang S., Ye Y., Zhu K., Yan L., Lu Y. та ін.
  ToolLLM: Facilitating Large Language Models to Master 16000+ Real-world APIs.
  \textit{Proc. 12th International Conference on Learning Representations (ICLR 2024)}.
  OpenReview.net, 2024.

\item Schick T., Dwivedi-Yu J., Dessì R., Raileanu R., Lomeli M., Hambro E.,
  Zettlemoyer L., Cancedda N., Scialom T.
  Toolformer: Language Models Can Teach Themselves to Use Tools.
  \textit{Advances in Neural Information Processing Systems 36 (NeurIPS 2023)}.
  Curran Associates, 2023.

\item Wang X., Chen Y., Yuan L., Zhang Y., Li Y., Peng H., Ji H.
  Executable Code Actions Elicit Better LLM Agents.
  \textit{Proc. 41st International Conference on Machine Learning (ICML 2024)}.
  PMLR, 2024.
  С.~50208--50232.

\item Wei A., Nie A., Teixeira T.~S.~F.~X., Yadav R., Lee W., Wang K., Aiken A.
  Improving Parallel Program Performance with LLM Optimizers via Agent--System Interfaces.
  \textit{Proc. 42nd International Conference on Machine Learning (ICML 2025)}.
  PMLR, 2025.
  arXiv:2410.15625.

\item Yao S., Zhao J., Yu D., Du N., Shafran I., Narasimhan K., Cao Y.
  ReAct: Synergizing Reasoning and Acting in Language Models.
  \textit{Proc. 11th International Conference on Learning Representations (ICLR 2023)}.
  OpenReview.net, 2023.

\end{citreflist}
