% sections/01-problem.tex --- Постановка наукової проблеми.

\citsection{The Compute Constraint in LLM Agent Toolchains}

Recent years have witnessed substantial progress in the capabilities of language-model agents along several dimensions: multi-step reasoning, programmatic code generation, retrieval-grounded answering, and the orchestration of long sequences of external tool invocations. The computational substrate on which such agents execute their actions, however, has remained essentially unchanged. The dominant architectural pattern, established by Toolformer and ReAct and adopted by all major contemporary agent frameworks, executes agent-generated code within a single-threaded sandbox, typically a Python interpreter, a code-interpreter container, or an analogous single-process runtime. Within this paradigm the model produces straight-line code, a single process executes it, and the agent's wall-clock latency coincides with the script's serial execution time.

The single-threaded execution model is adequate for agent behaviour built around retrieval, parsing, calculation, and lightweight scripting. It becomes inadequate, however, for an expanding class of practically relevant workloads whose total computational work scales with sample count or problem size. Representative examples include Monte Carlo estimation, hyperparameter exploration, combinatorial optimisation, large-scale simulation, and statistical bootstrap. When executed within a single process, such workloads either exceed the available time budget or are forced into severe undersampling that degrades the quality of the obtained results; in either case the reasoning capacity of the underlying model is constrained not by its inferential ability but by the available compute resources.

A direct alternative is to provide the agent with access to a parallel runtime. This approach, however, encounters a documented difficulty: the generation of correct parallel code remains an unsolved subproblem for contemporary LLMs. Empirical evaluations report substantial accuracy degradation in MPI, OpenMP, and CUDA generation tasks relative to equivalent serial implementations, particularly for non-trivial memory-access patterns. Models capable of producing fluent serial implementations frequently fail to express the same algorithms when explicit synchronisation primitives are required.

The two requirements -- parallel execution and synchronous-style code generation -- need not be mutually exclusive. A distributed runtime can absorb parallelism transparently, accepting per-worker logic written in straight-line code while assuming responsibility for fan-out, fan-in, and result aggregation. The PARCS framework constitutes such a runtime: a Kubernetes-based cluster of daemon workers that executes user-supplied modules at runtime. The remaining design problem is the agent-facing interface.

The interface considered in the present work is the Model Context Protocol (MCP), a recently introduced standard for agent--tool interaction. Tools advertised through MCP are usable by any compliant agent without bespoke integration. Positioning a distributed-computing cluster behind MCP and exposing a compact set of tools aligned with the agent's natural step-wise decomposition allows agent-generated C\# code to retain a single-threaded coding paradigm while parallel execution is delivered transparently by the platform. The research problem addressed in this paper is therefore the design of an interface between a language-model agent and a distributed parallel computing cluster that preserves the synchronous coding paradigm at the agent's surface while providing cluster-scale computational throughput beneath it.
