% sections/02-related-work.tex --- Аналіз досліджень.
% Citation key map (draft.md numeric -> bib key):
%   [1]=schick2023toolformer  [2]=yao2023react  [3]=qin2024toolllm
%   [4]=patil2023gorilla      [5]=ma2024sciagent [6]=mialon2023augmented
%   [7]=wang2024codeact       [8]=kim2024llmcompiler
%   [9]=nichols2024pareval    [10]=nichols2024hpccoderv2
%   [11]=xu2025llmhpc         [12]=he2025legion
%   [13]=anthropic2024mcp     [14]=hou2025mcp  [15]=luo2025mcpsmelly
%   [16]=bohusevych2026parcs  [17]=bohusevych2024tsp

\citsection{Analysis of related research}

\textit{Tool-augmented language-model agents.}\quad
Toolformer~\cite{schick2023toolformer} trained a model to insert API calls into its own output without supervised labels; ReAct~\cite{yao2023react} added explicit reasoning traces interleaved with observations from search and database tools. Later work extended both the scope and the toolset: ToolLLM~\cite{qin2024toolllm} covered 16\,000 REST endpoints, Gorilla~\cite{patil2023gorilla} targeted hallucinated API signatures in machine-learning framework calls, and SciAgent~\cite{ma2024sciagent} brought the same approach to scientific domains spanning mathematics, physics, and finance. Mialon et al.~\cite{mialon2023augmented} survey this family under the label of augmented language models. All of these systems share one structural constraint: every tool invocation is a local synchronous function call in which the server receives arguments, returns a result, and terminates, while the agent's host process remains the only site of computation.

CodeAct~\cite{wang2024codeact} generalises the action space to arbitrary Python code executed within a persistent code-interpreter sandbox, enabling stateful execution across reasoning turns and access to numerical libraries and file I/O. The substrate, however, remains a single interpreter, and the agent's wall-clock cost is determined by the script's serial execution time. LLMCompiler~\cite{kim2024llmcompiler} introduces a form of parallelism by analysing a query into a dependency graph and dispatching independent calls to existing tools concurrently, yielding latency reductions of up to a factor of 3.7 on multi-hop question-answering tasks. The parallelism it exploits, however, exists at the level of tool calls rather than within a single computation, and the system neither provisions compute resources nor compiles new code.

\textit{Language models and parallel code generation.}\quad
A separate line of research evaluates the capability of language models to generate parallel programs. ParEval~\cite{nichols2024pareval} benchmarks code generation in MPI, OpenMP, CUDA, and Kokkos across 420 problems and reports a marked decrease in correctness relative to equivalent serial implementations, with the gap widening as problem complexity increases. HPC-Coder-V2~\cite{nichols2024hpccoderv2} reduces this gap through fine-tuning on a synthetic parallel-code corpus, while a recent evaluation of frontier models on classical high-performance computing kernels~\cite{xu2025llmhpc} confirms that non-trivial memory-access patterns remain a persistent source of error. He et al.~\cite{he2025legion} employ language models to optimise mappers in the Legion distributed runtime; this constitutes an effective specialisation, but it presupposes the existence of a parallel program rather than its generation.

The cumulative finding of this line of work is that requiring a language model to express parallelism explicitly degrades the quality of its output. A model that produces a clean serial loop frequently fails to render the same loop correctly when synchronisation primitives such as \texttt{MPI\_Bcast}, \texttt{\#pragma omp parallel for}, or a CUDA kernel launch are involved.

\textit{The Model Context Protocol.}\quad
The Model Context Protocol, introduced by Anthropic in 2024~\cite{anthropic2024mcp}, specifies a standard JSON-RPC interface for agent--tool integration. The survey of Hou et al.~\cite{hou2025mcp} catalogues the early MCP ecosystem and identifies security concerns including prompt injection and tool poisoning, but reports no published work that uses MCP to expose distributed computing infrastructure. A complementary study of public MCP tool descriptions~\cite{luo2025mcpsmelly} established that more than half of the surveyed tools provide inadequate documentation of purpose or parameters; this finding motivates the verbosity of the tool-description design adopted in the following section.

\textit{The PARCS framework.}\quad
PARCS (Parallel Asynchronous Recursive Control Space) is a distributed computing platform whose computational model originates in early-1980s research on non-von~Neumann architectures~\cite{anisimov2017parcs}. Its core abstraction is the \textit{control space}: a finite, dynamically extensible graph of \textit{points} --- addressable locations where executable algorithmic modules operate --- connected by bidirectional \textit{channels}. Topologies such as trees, rings, and hypercubes are formed and modified at runtime; modules communicate asynchronously via message-passing and may spawn new points and channels during execution, giving the model its recursive character. The platform has been implemented across eight language extensions spanning PASCAL, C, Fortran, Java, CUDA, Python, and C\#, with earlier cloud deployments targeting Amazon~EC2 and Microsoft~Azure. The present work uses the current incarnation --- PARCS-C\#, a .NET\,10 Kubernetes-native implementation that packages daemon workers as containerised pods orchestrated by KEDA and exposes job management through a REST Host~API~\cite{bohusevych2026parcs}. A previous publication~\cite{bohusevych2024tsp} extended this deployment with KEDA-driven autoscaling and evaluated it on a distributed island-model genetic algorithm for the Travelling Salesman Problem, showing 20--48\,\% shorter routes than sequential execution on the same time budget with 13--25\,\% communication overhead. That result establishes the viability of PARCS as a parallel runtime; it does not address the agent-facing interface that is the subject of this paper.

\textit{Identified gap and research aim.}\quad
Three observations follow from the literature reviewed above. First, every published agent framework executes agent-generated code within a single-threaded substrate, which limits the effective compute envelope available to the agent. Second, attempts to obtain parallelism by requesting the agent to generate parallel code conflict with the documented deficiencies of LLMs in this task. Third, no prior work employs MCP as the interface between an agent and a distributed compute runtime, despite MCP being well-suited to the role.

The aim of the present work is therefore to design and describe an architectural pattern in which a language-model agent harnesses a Kubernetes-based PARCS cluster through Model Context Protocol tools by submitting per-worker logic in straight-line C\# at runtime, without itself producing parallel code. The remainder of the paper presents this pattern, the protocol surface, and a Monte Carlo Value-at-Risk case study illustrating the full execution lifecycle. Quantitative evaluation is reserved for a separate publication.
